Distributional Trees and Forests

### 2017-10-11 pkg/disttree/inst/draft/draft.Rnw
added PIT-histograms and QQR-plots
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@1651 255f4249-2263-4417-b1ea-8498779e27a7

\SweaveOpts{engine=R, eps=FALSE, keep.source=TRUE}
<<preliminaries, echo=FALSE, results=hide>>=
options(prompt = "R> ", continue = "+  ", useFancyQuotes = FALSE, width = 70)
library("ggplot2")
theme_set(theme_bw(base_size = 18))
library("disttree")
library("Formula")  # FIX ME: should be imported in disttree
library("gamlss")
library("randomForest")
library("lattice")
library("crch")
library("latex2exp")
library("parallel")
library("gamlss.cens")
gen.cens(NO, type = "left")
library("mgcv")
library("MASS")

@@

\begin{figure}[t!]
%\begin{subfigure}{0.5\textwidth}
<<rain_axams_varim_margeffects1_mu, fig=TRUE, echo=FALSE>>=
par(mar = c(2,10,2,2))
barplot(sort(vimp_crps, decreasing = TRUE)[1:10], horiz = TRUE, las = 1, axes = FALSE)
@
%\end{subfigure}
%\hspace{-1cm}
%\begin{subfigure}{0.5\textwidth}
%<<rain_axams_varim_margeffects1_sigma, fig=TRUE, echo=FALSE>>=
%plot(x = sp1[,id_mi[1]], y = sp1[, "sigma"], type = "l", xlab = names_mi[1], ylab = %"sigma")
%@
%\end{subfigure}
\caption{\label{varimp} Variable imortance: top 10 list for the forest model}
\end{figure}

>>>>

\SweaveOpts{engine=R, eps=FALSE, keep.source=TRUE}
<<preliminaries, echo=FALSE, results=hide>>=
options(prompt = "R> ", continue = "+  ", useFancyQuotes = FALSE, width = 70)
library("ggplot2")
theme_set(theme_bw(base_size = 18))
library("disttree")
library("Formula")  # FIX ME: should be imported in disttree
library("gamlss")
library("randomForest")
library("lattice")
library("crch")
library("latex2exp")
library("parallel")
library("gamlss.cens")
gen.cens(NO, type = "left")
library("mgcv")
library("MASS")
library("countreg")

@@

\begin{figure}[t!]
%\begin{subfigure}{0.5\textwidth}
<<rain_axams_varim_margeffects1_mu, fig=TRUE, echo=FALSE>>=
par(mar = c(2,10,2,2))
barplot(sort(vimp_crps, decreasing = TRUE)[1:10], horiz = TRUE, las = 1, cex.names = 0.8, axes = FALSE)
@
%\end{subfigure}
%\hspace{-1cm}
%\begin{subfigure}{0.5\textwidth}
%<<rain_axams_varim_margeffects1_sigma, fig=TRUE, echo=FALSE>>=
%plot(x = sp1[,id_mi[1]], y = sp1[, "sigma"], type = "l", xlab = names_mi[1], ylab = %"sigma")
%@
%\end{subfigure}
\caption{\label{varimp} Variable imortance: top 10 list for the forest model}
\end{figure}

<<pit_qqr, echo=FALSE, eval=TRUE>>=

## pit histograms and qqr-plots

## get predicted parameter
# for learndata and testdata
{

  ## disttree
  # in sample
  pdt <- predict(res$dt, type = "parameter")
  dt_mu_l <- pdt$mu 
  dt_sigma_l <- pdt$sigma 
  pit_dt_l <- cbind(0, pnorm(res$learndata[,"robs"], mean = dt_mu_l, sd = dt_sigma_l))
  pit_dt_l[res$learndata[,"robs"]>0, 1] <- pit_dt_l[res$learndata[,"robs"]>0, 2]

  # out of sample
  pdt <- predict(res$dt, type = "parameter", newdata = res$testdata)
  dt_mu_t <- pdt$mu 
  dt_sigma_t <- pdt$sigma 
  pit_dt_t <- cbind(0, pnorm(res$testdata[,"robs"], mean = dt_mu_t, sd = dt_sigma_t))
  pit_dt_t[res$testdata[,"robs"]>0, 1] <- pit_dt_t[res$testdata[,"robs"]>0, 2]

  ## distforest
  # in sample
  pdf <- predict(res$df, type = "parameter")
  df_mu_l <- pdf$mu
  df_sigma_l <- pdf$sigma
  pit_df_l <- cbind(0, pnorm(res$learndata[,"robs"], mean = df_mu_l, sd = df_sigma_l))
  pit_df_l[res$learndata[,"robs"]>0, 1] <- pit_df_l[res$learndata[,"robs"]>0, 2]

  # out of sample
  pdf <- predict(res$df, type = "parameter", newdata = res$testdata)
  df_mu_t <- pdf$mu
  df_sigma_t <- pdf$sigma
  pit_df_t <- cbind(0, pnorm(res$testdata[,"robs"], mean = df_mu_t, sd = df_sigma_t))
  pit_df_t[res$testdata[,"robs"]>0, 1] <- pit_df_t[res$testdata[,"robs"]>0, 2]

  ## gamlss
  # in sample
  g_learndata <- res$learndata
  g_learndata$robs <- Surv(g_learndata$robs, g_learndata$robs>0, type="left")
  g_mu_l <- predict(res$g, what = "mu", type = "response")
  g_sigma_l <- predict(res$g, what = "sigma", type = "response")
  pit_g_l <- cbind(0, pnorm(res$learndata[,"robs"], mean = g_mu_l, sd = g_sigma_l))
  pit_g_l[res$learndata[,"robs"]>0, 1] <- pit_g_l[res$learndata[,"robs"]>0, 2]

  # out of sample
  g_mu_t <- predict(res$g, what = "mu", type = "response", newdata = res$testdata)
  g_sigma_t <- predict(res$g, what = "sigma", type = "response", newdata = res$testdata)
  pit_g_t <- cbind(0, pnorm(res$testdata[,"robs"], mean = g_mu_t, sd = g_sigma_t))
  pit_g_t[res$testdata[,"robs"]>0, 1] <- pit_g_t[res$testdata[,"robs"]>0, 2]

  #gamboostLSS
  pgb <- predict(res$gb, parameter = c("mu","sigma"), type = "response")
  gb_mu_l <- pgb$mu
  gb_sigma_l <- pgb$sigma
  pit_gb_l <- cbind(0, pnorm(res$learndata[,"robs"], mean = gb_mu_l, sd = gb_sigma_l))
  pit_gb_l[res$learndata[,"robs"]>0, 1] <- pit_gb_l[res$learndata[,"robs"]>0, 2]

  # out of sample
  pgb <- predict(res$gb, parameter = c("mu","sigma"), type = "response", newdata = res$testdata)
  gb_mu_t <- pgb$mu
  gb_sigma_t <- pgb$sigma
  pit_gb_t <- cbind(0, pnorm(res$testdata[,"robs"], mean = gb_mu_t, sd = gb_sigma_t))
  pit_gb_t[res$testdata[,"robs"]>0, 1] <- pit_gb_t[res$testdata[,"robs"]>0, 2]

  #EMOS
  ml_mu_l <- predict(res$ml, type = "location")     # returns parameter on response scale
  ml_sigma_l <- predict(res$ml, type = "scale")
  pit_ml_l <- cbind(0, pnorm(res$learndata[,"robs"], mean = ml_mu_l, sd = ml_sigma_l))
  pit_ml_l[res$learndata[,"robs"]>0, 1] <- pit_ml_l[res$learndata[,"robs"]>0, 2]

  # out of sample
  ml_mu_t <- predict(res$ml, type = "location", newdata = res$testdata)     # returns parameter on response scale
  ml_sigma_t <- predict(res$ml, type = "scale", newdata = res$testdata)
  pit_ml_t <- cbind(0, pnorm(res$testdata[,"robs"], mean = ml_mu_t, sd = ml_sigma_t))
  pit_ml_t[res$testdata[,"robs"]>0, 1] <- pit_ml_t[res$testdata[,"robs"]>0, 2]

\begin{figure}[t!]
%\begin{subfigure}{0.5\textwidth}
<<rain_axams_pit_insample, fig=TRUE, echo=FALSE>>=

# in sample
set.seed(7)
par(mfrow = c(2,2))
pithist(pit_df_l, nsim = 1000, breaks = seq(0, 1, length.out = 9))
pithist(pit_g_l, nsim = 1000, breaks = seq(0, 1, length.out = 9))
pithist(pit_gb_l, nsim = 1000, breaks = seq(0, 1, length.out = 9))
pithist(pit_ml_l, nsim = 1000, breaks = seq(0, 1, length.out = 9))

@
\caption{\label{pit_insample} PIT-Histograms for station Axams (in sample)}
\end{figure}

\begin{figure}[t!]
%\begin{subfigure}{0.5\textwidth}
<<rain_axams_pit_oosample, fig=TRUE, echo=FALSE>>=

# out of sample
set.seed(7)

par(mfrow = c(2,2))
pithist(pit_df_t, nsim = 1000, breaks = seq(0, 1, length.out = 9))
pithist(pit_g_t, nsim = 1000, breaks = seq(0, 1, length.out = 9))
pithist(pit_gb_t, nsim = 1000, breaks = seq(0, 1, length.out = 9))
pithist(pit_ml_t, nsim = 1000, breaks = seq(0, 1, length.out = 9))

@
\caption{\label{pit_oosample} PIT-Histograms for station Axams (out of sample)}
\end{figure}

\begin{figure}[t!]
%\begin{subfigure}{0.5\textwidth}
<<rain_axams_qqr_insample, fig=TRUE, echo=FALSE>>=
# in sample
set.seed(7)
par(mfrow = c(2,2))
set.seed(1)
qqrplot(pit_df_l, nsim = 1, main = "distforest (in bag)")
qqrplot(pit_g_l, nsim = 1, main = "gamlss (in bag)")
qqrplot(pit_gb_l, nsim = 1, main = "gamboostLSS (in bag)")
qqrplot(pit_ml_l, nsim = 1, main = "EMOS (in bag)")

@
\caption{\label{qqr_insample} QQR-Plot for station Axams (in sample)}
\end{figure}

\begin{figure}[t!]
%\begin{subfigure}{0.5\textwidth}
<<rain_axams_qqr_oosample, fig=TRUE, echo=FALSE>>=

# out of sample
set.seed(7)
par(mfrow = c(2,2))
set.seed(1)
qqrplot(pit_df_t, nsim = 1, main = "distforest (out of bag)")
qqrplot(pit_g_t, nsim = 1, main = "gamlss (out of bag)")
qqrplot(pit_gb_t, nsim = 1, main = "gamboostLSS (out of bag)")
qqrplot(pit_ml_t, nsim = 1, main = "EMOS (out of bag)")

@
\caption{\label{qqr_oosample} QQR-Plot for station Axams (out of sample)}
\end{figure}
